---
title: "Basic R Notebook"
output:
  html_notebook: default
  pdf_document: default
---

Preliminaries

Returns the current working directory
```{r}
getwd()
```
Provides help on a specific fucntion
```{r}
help(getwd)
?getwd
?setwd
```
Sets the current working directory
```{r}
setwd("C:/Users/natarajan_karthik/Desktop/AE")
```
Lists the files in a directory
```{r}
dir()
```
Lists objects stored in the R workspace
```{r}
ls()
```
Assign a number to a variable
```{r}
x <- 50
x
ls()
```
Alternative to assign a number to a variable
```{r}
y = 50
y
```
Common functions - exponential, inverse, power, add
```{r}
exp(x)
1/x
x^3
y <- x+6
```
Remove a variable from the workspace
```{r}
rm(y)
ls()
```
Numbers and vectors

Concatenates (combines) numbers to form a vector
```{r}
x <- c(1,-2,3,5,pi)
x
```
Accessing specific elements of the vector
```{r}
x[2]
```
Applying operations to the vector - term by term inverse, concatenate vectors, exponentiation
```{r}
1/x
y <- c(x,0,x)
exp(y)
y
```
You can overload the sum operator by recycling the shorter vector - mathematically adding vectors of different sizes are not permitted 
```{r}
y+x
```
Finding the maximum and minimum elements and identifying the location (index) of the first max and all max
```{r}
max(x)
min(x)
?which.max
which.max(x)
which(x==max(x))
z <- c(x,5)
which.max(z)
which(z==max(z))
```
Other operations - sum, product, mean, variance, standard deviation
```{r}
sum(x)
prod(x)
mean(x)
var(x)
sd(x)
```
Maxium entry in a concatenated vector
```{r}
max(x,y)
```
Parallel maximum returns a vector of length equal to the longest argument that contains in each element, the largest element in that position of any of the vector
```{r}
?pmax
pmax(x,y)
```
For a vector, it provides a six number summary including min, max, mean, 1st quartile, median and 3rd quartile. This can be used with other objects too.
```{r}
?summary
summary(x)
```
Remove all variables from the workspace
```{r}
rm(list=ls())
```
Differences in the assignment using <- and ==. Try exp(a=1:5) and exp(a<-1:5)
```{r}
exp(a=1:5)
a
```
```{r}
exp(a<-1:5)
a
```
Generating vectors using a variety of commands
```{r}
x <- -4:4
?seq
seq(-4,4,0.2)
?rep
x
rep(x,times=2)
rep(x,each=2)
```
Returns logical vector based on the check
```{r}
x > 1
```
Dealing with missing entries 
```{r}
?is.na
is.na(x)
y <- c(x,NA)
y
is.na(y)
mean(y)
?mean
mean(y,na.rm=TRUE)
```
R stores objects in a variety of classes - numeric, integer, character, logical, list, matrix, dataframe and has logical overriding operations when you convert from one class to another
```{r}
z <- 0:9
z
class(z)
z1 <- c("a","b")
z1
class(z1)
w <- as.character(z)
w
class(w)
as.integer(w)
as.logical(c(5,0))
w <- c(1,"a")
w
T
F
```
Factor 
```{r}
x <- 1:7
class(x)
y <- factor(x)
y
levels(y)
summary(y)
x
income <- c(500,1200,4000,2300,2300,1234,1345)
```
Computes the mean of subvectors in income from the factors in x
```{r}
income <- c(500,1200,4000,2300,2300,1234,1345)
tapply(income,x,mean)
x
```
Matrix, array - operations
```{r}
?matrix
r <- matrix(c(3,4,5,6,7,8),nrow=3,ncol=2)
r
dim(r)
r[2,2]
r[5]
r[1,]
class(r)
?array
s <- array(c(3,4,5,6,7,8),c(3,2))
s
class(s)
z <- 1:50
z
?dim
dim(z) <- c(5,2,5)
z[5,2,5]
z[5,2,1:5]
diag(10)
?cbind
cbind(c(1,2,3),c(4,5,6))
rbind(c(1,2,3),c(4,5,6))
z <- matrix(c(5,7,9,6,3,4),nrow=3,ncol=2)
y <- matrix(c(1,3,0,9,5,-1),nrow=3,ncol=2)
z*y
z
y
help("%*%")
z %*% t(y)
a <- array(c(2,1,-1,2),c(2,2))
b <- c(4,4)
a
b
?solve
solve(a,b)
solve(a)
eigen(a)
E <- eigen(a)
E$value
E$vectors
```
List consists of an ordered collection of objects that can be of different or the same type.
```{r}
c(1,2,"a")
?list
karthik <- list(age=39,sex="M",child.ages=c(3,3))
class(karthik)
karthik
kim <- list(age=30,sex="F",child.ages=NA)
kim
karthik$age
karthik$child.ages
karthik$child.ages[1]
karthik[1]
karthik[2]
karthik[3]
class(karthik[1])
karthik[[1]]
class(karthik[[1]])
kk <- c(karthik,kim)
kk
kk$age
kk[1]
kk[3]
kk[4]
```
Dataframes are a tightly coupled collection of variables that share many of the properties of matrices and lists and is the fundamental data structure that will be used in most of this course.
```{r}
?data.frame
A <- data.frame(names=c("karthik","kim"),ages=c(39,30),children=c(2,0))
class(A)
A$names
class(A$names)
A$ages
class(A$ages)
A$spouse <- c("sam","jim")
```

