---
title: "R Notebook"
output: html_notebook
---

*****************************
Preparation

Reading in Files
```{r}
software <- read.csv("software.csv")
breast <- read.csv("breastcancer.csv")

str(software)
str(breast)

```


*****************************
Question 1

1.1 The amount spent by customers in purchasign from the mailed catalog is normally distrubted

```{r}
hist(software$spending, breaks = seq(0,2000,100))



```

1.2 Avg $ and variance of $ spent by males and females respectively

```{r}

mean(software$spending[software$gender==1]) #males
sd(software$spending[software$gender==1])^2
var(software$spending[software$gender==1])
mean(software$spending[software$gender==0]) #females
var(software$spending[software$gender==0])


```

1.3 t-test

```{r}
avgspendM <- software$spending[software$gender==1]
avgspendF <- software$spending[software$gender==0]
t.test(avgspendM,avgspendF)


```


1.4 LM model 
```{r}
#software$partition
softtrain <- software[software$partition=='t',]
softtest <- software[software$partition=='test',]
softval <- software[software$partition=='v',]


lm14 <- lm(data=softtrain,spending~freq+last_update+first_update+web_order+address_res+address_us)
summary(lm14)


```

1.5 ans qn using model
A) cust move from non-res addr to resi addr (0 to 1)
B) cust numb of trans in a year decreases by 1 unit

on avg, which of the two, will the company be worse off in terms of revenue (spending)

> A, cos lose more $$


1.6 build smaller model only sig var but leave intercept inside
compare adj R, which you prefer


```{r}
lm16 <- lm(data=softtrain,spending~freq+address_res)
summary(lm16)


```

1.7 leaps to do subset sel
backwards 
which is drop first?
comment on how this result relates to pvalue you est for FULL model

```{r}
library(leaps)
model17 <- regsubsets(spending~freq+last_update+first_update+web_order+address_res+address_us,softtrain,nvmax=10,method='backward')
summary(model17)

```

1.8 use adj r2 to pick best model
which var are included

```{r}
plot18 <- summary(model17)$adjr2
plot(plot18)

maxRsquare <- which.max(plot18)
maxRsquare
plot18[maxRsquare]

summary(model17)$which
summary(model17)$which[maxRsquare,]



```

1.9 use val set to choose btwn 3 models
for all 3 what is the sse in val set


```{r}
ModelPredFull <- predict(lm14, newdata = softval)
ModelPredSmall <- predict(lm16, newdata = softval)

lm17 <- lm(data=softtrain,spending~freq+last_update+address_res+address_us)
ModelPredSubset <- predict(lm17, newdata = softval)

sseFULL <- sum( (ModelPredFull - softval$spending)^2 )
sseFULL

sseSMALL <- sum( (ModelPredSmall - softval$spending)^2 )
sseSMALL

sseSUBSET <- sum( (ModelPredSubset - softval$spending)^2 )
sseSUBSET



```
choose modelSMALL


1.10 evaluate on test set
what is sse
are they consistent?


```{r}
ModelPredFull <- predict(lm14, newdata = softtest)
ModelPredSmall <- predict(lm16, newdata = softtest)
ModelPredSubset <- predict(lm17, newdata = softtest)

sseFULL <- sum( (ModelPredFull - softtest$spending)^2 )
sseFULL

sseSMALL <- sum( (ModelPredSmall - softtest$spending)^2 )
sseSMALL

sseSUBSET <- sum( (ModelPredSubset - softtest$spending)^2 )
sseSUBSET



```

Pick small still.



**********************
Question 2

2.1 Which var has most numb of missing entries
```{r}
str(breast)

summary(breast)

```
Density most missing  @ 76 


2.2 got balance? (check severity levels)
```{r}
table(breast$severity)

516/(516+445)
445/(516+445)
```
Yes


2.3 Drop all NA entries using na.omit
split into test and training 60% of data into training
set seed to 1000

write the est eqn that predicts sev using age shape margin and dens and instercept


```{r}
newbreast<- na.omit(breast)
summary(newbreast)

set.seed(1000)
library(caTools)
breastsplit <- sample.split(newbreast$severity, SplitRatio = 0.6)
breasttrain <- subset(newbreast, breastsplit == TRUE)
breasttest <- subset(newbreast, breastsplit == FALSE)

glm23 <- glm(data=breasttrain,severity~age+shape+margin+density,family=binomial)
summary(glm23)
```

2.4 whats the odd ratio of age
exp(0.05989)


2.5 what is LL of model on test set ??????????
what is acc of model thres 0.5
```{r}

BreastModelPredict <- predict(glm23, newdata = breasttest,type="response")
summary(BreastModelPredict)

-0.5*glm23$deviance

table(as.integer(BreastModelPredict >= 0.5),breasttest$severity)

(127+139)/(127+139+22+44)
```


2.6 use only stat sig pred var include intercept and rebuild
what is LL on test set and acc?

```{r}

glm26 <- glm(data=breasttrain,severity~age+shape+margin,family=binomial)
summary(glm26)


-0.5*glm26$deviance

BreastModelPredict2 <- predict(glm26, newdata = breasttest,type="response")
table(as.integer(BreastModelPredict2 >= 0.5),breasttest$severity)

(126+141)/(125+141+20+45)


```

2.7 calc partial area under ROC far FP rates which is 0.7 or larger (use model in 2.6)


```{r}
library(ROCR)
ROCR_Predict <- prediction(BreastModelPredict2, breasttest$severity)

performance(ROCR_Predict, measure = "auc", fpr.stop=0.7)


```

2.8 probit??
what is LL on test set and acc?
```{r}
glm28 <- glm(data=breasttrain,severity~age+shape+margin,family=binomial(link="probit"))
summary(glm28)


-0.5*glm28$deviance

BreastModelPredict2 <- predict(glm28, newdata = breasttest,type="response")
table(as.integer(BreastModelPredict2 >= 0.5),breasttest$severity)

(126+140)/(125+140+21+45)


```

2.9 which is better?
Logistic


2.10 

```{r}
newerbreast <- newbreast
newerbreast$severity <- as.integer(newbreast$physician>=4)

glm210 <- glm(data=newerbreast,severity~age+shape+margin,family=binomial)
summary(glm210)

BreastModelPredict3 <- predict(glm210, newdata = newerbreast,type="response")
table(as.integer(BreastModelPredict3 >= 0.5),newerbreast$severity)

table(newerbreast$severity)
table(newbreast$severity)

794-403

```



3.1
```{r}
seq(4,100,by=2)
seq(6,100,by=2)
cumprod(c(4,6,8)
cumprod(seq(4,100,by=2))



```



```{r}


0.2*0.4 + 0.2
0.4*0.4 + 0.4
```

karthik_natarajan@sutd.edu.sg

```{r}




```



```{r}




```



```{r}




```



```{r}




```



```{r}




```



```{r}




```



```{r}




```



```{r}




```



```{r}




```



```{r}




```



```{r}




```



```{r}




```



























